---
title: "p8105_hw3_jl6321.Rmd"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1 

(corrected answer after checking the posted solution)

Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

There are `r nrow(instacart)` rows and `r ncol(instacart)` columns, where fresh vegetables and fresh fruits holding the most ordered items.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Make a plot that shows the number of items ordered in each aisle. 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. 

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

## Problem 2

Read and tidy the data. 

```{r Problem2-1}
accel_df = 
  read_csv('./Data/accel_data.csv') %>% 
  pivot_longer( 
    activity.1:activity.1440, 
    names_to = "minute", 
    names_prefix = "activity.", 
    values_to = "activity_counts" 
  ) %>% 
  mutate(if_weekend = (day %in% c('Saturday', 'Sunday'))) %>% 
  relocate(if_weekend, .after = day_id)
```

Variables in this dataset include `r colnames(accel_df)`. There are `r nrow(accel_df)` observations in total. 

Aggregate across minutes for each day and create a table. 

```{r Problem2-2}
accel_aggregate = accel_df %>%
  group_by(week, day) %>%
  summarise(total_activity = sum(activity_counts)) %>%
  pivot_wider(
    names_from = "day",
    values_from = "total_activity"
  ) 
accel_aggregated = accel_aggregate[, c(1, 3, 7, 8, 6, 2, 4, 5)]
accel_aggregated
```

There does not seem to be any apparent trend based on the table besides that the total activities on the 4th and 5th Saturday tend to be low. 

Make a single-panel plot. 

```{r Problem2-3}
accel_df %>% 
  mutate(minute = as.numeric(minute)) %>% 
  group_by(day, minute) %>% 
  summarize(mean_activity_counts = mean(activity_counts)) %>% 
  ggplot(aes(x = minute, y = mean_activity_counts, color = day)) + 
  geom_smooth(se = FALSE) + 
  labs( 
    title = "Average Activity Time Across the Day", 
    x = "Hour", 
    y = "Average Activity Time", 
    caption = "Data from accel.csv dataset") + 
  scale_x_continuous(breaks = c(0:24) * 60, labels = c(0:24))
```

According to the graph, we can conclude that there tends to be most activity on average around 9pm on Friday and 11am on Sunday. The average activity seems to be low from 11pm to 5am the next day. 

## Problem 3 

```{r Problem3-0}
data("ny_noaa")
missing_data = sum(is.na(ny_noaa))
missing_rate = missing_data / (nrow(ny_noaa) * ncol(ny_noaa))
```

This dataset consists of `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. The variables include `r colnames(ny_noaa)`. The dataset possesses a missing data rate of `r missing_rate`, which may be an issue because the percentage is greater than 10%. 

Clean the dataset. 

```{r Problem3-1-1}
ny_noaa_tidy = 
  ny_noaa %>%
  separate(date, into = c("year", "month", "day"), sep = "-") %>%
  mutate(
    prcp = as.double(prcp) / 10, 
    tmax = as.double(tmax) / 10, 
    tmin = as.double(tmin) / 10
  )
```

Find the most commonly observed snowfall value. 

```{r Problem3-1-2}
snowfall = 
  ny_noaa_tidy %>% 
  group_by(snow) %>%
  summarise(n_obs = n()) %>%
  arrange(desc(n_obs)) 
```

The most commonly observed value for snowfall is `r snowfall$snow[1]`mm. 
Make a two-panel plot showing the average max temperature in January and in July in each station across years. 

```{r Problem3-2}
ny_noaa_tidy %>% 
  filter(month == "01" | month == "07") %>% 
  group_by(id, year, month) %>% 
  summarize(tmax_mean = mean(tmax, rm.na = TRUE)) %>% 
  ggplot(aes(x = as.integer(year), y = tmax_mean, color = id)) + 
  geom_point() + 
  geom_line(alpha = 0.3) +
  facet_grid(. ~ month) + 
  labs(
    title = "Mean Max Temperature in January and July Across Years", 
    x = "Year", 
    y = "Mean Max Temperature in C"
  ) + 
  theme(legend.position = "none")
```

The overall average max temperature in July is higher than that in January. Compared with July, the average max temperature in January seems to have a greater range. Some outliers exist, including one datapoint in January of 1982 and one datapoint in July of 1988. 

Make a two-panel plot showing: 
(i) tmax vs tmin for the full dataset 
(ii) the distribution of snowfall values > 0 and < 100 separately by year.

```{r Problem3-3}
tmax_vs_tmin = 
  ny_noaa_tidy %>% 
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex(alpha = 0.9) + 
  labs(
    title = "Max Temperature vs. Min Temperature", 
    x = "Max Temperature in C", 
    y = "Max Temperature in C", 
    caption = "Data from the ny_noaa dataset") 

snowfall_distribution = 
  ny_noaa_tidy %>% 
  filter(snow > 0 & snow < 100) %>% 
  ggplot(aes(x = snow, y = year)) + 
  geom_density_ridges(scale = 0.9, fill = 'chartreuse4') + 
  labs(
    title = "The Distribution of Snowfall by Year", 
    x = "Snowfall in mm", 
    y = "Year", 
    caption = "Data from the ny_noaa dataset"
  ) 

tmax_vs_tmin + snowfall_distribution 
```

