---
title: "p8105_hw3_jl6321.Rmd"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 2

Read and tidy the data. 

```{r Problem2-1}
accel_df = 
  read_csv('./Data/accel_data.csv') %>% 
  pivot_longer( 
    activity.1:activity.1440, 
    names_to = "minute", 
    names_prefix = "activity.", 
    values_to = "activity_counts" 
  ) %>% 
  mutate(if_weekend = (day %in% c('Saturday', 'Sunday'))) %>% 
  relocate(if_weekend, .after = day_id)
```

Variables in this dataset include `r colnames(accel_df)`. There are `r nrow(accel_df)` observations in total. 

Aggregate across minutes for each day and create a table. 

```{r Problem2-2}
accel_aggregate = accel_df %>%
  group_by(week, day) %>%
  summarise(total_activity = sum(activity_counts)) %>%
  pivot_wider(
    names_from = "day",
    values_from = "total_activity"
  ) 
accel_aggregated = accel_aggregate[, c(1, 3, 7, 8, 6, 2, 4, 5)]
accel_aggregated
```

There does not seem to be any apparent trend based on the table besides that the total activities on the 4th and 5th Saturday tend to be low. 

Make a single-panel plot. 

```{r Problem2-3}
accel_df %>% 
  mutate(minute = as.numeric(minute)) %>% 
  group_by(day, minute) %>% 
  summarize(mean_activity_counts = mean(activity_counts)) %>% 
  ggplot(aes(x = minute, y = mean_activity_counts, color = day)) + 
  geom_smooth(se = FALSE) + 
  labs( 
    title = "Average Activity Time Across the Day", 
    x = "Hour", 
    y = "Average Activity Time", 
    caption = "Data from accel.csv dataset") + 
  scale_x_continuous(breaks = c(0:24) * 60, labels = c(0:24))
```

According to the graph, we can conclude that there tends to be most activity on average around 9pm on Friday and 11am on Sunday. The average activity seems to be low from 11pm to 5am the next day. 

## Problem 3 

```{r}
data("ny_noaa")
missing_data = sum(is.na(ny_noaa))
missing_rate = missing_data / (nrow(ny_noaa) * ncol(ny_noaa))
```

This dataset consists of `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. The variables include `r colnames(ny_noaa)`. The dataset possesses a missing data rate of `r missing_rate`, which may be an issue because the percentage is greater than 10%. 


